Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
edits to 9.1

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the datacenter spine role based on router templates from a large cloud provider (\IE, macros over gobbledygook).}
  \label{fig:templates}
  \vspace{-.8em}
\end{figure}

@@

%While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they provide no guarantees
As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

\sysname takes as input the network's abstract topology and its policy. The policy has two components: the routing policy describes how traffic should flow, and the fault-tolerance policy describes how many simultaneous failures the network can withstand without losing connectivity.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the datacenter:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the variable $\CD{\$LP}$ and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the datacenter) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each datacenter level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the datacenter through a peer.

@@

The first step for the \sysname compiler is to build the product graph (PG), a data structure that is amenable to joint analysis of the topology and routing policy. As a precursor to this step, we convert the routing policy syntax in \S\ref{sec:overview} to a core language based on regular expressions, shown in Figure~\ref{fig:syntax}.
%
A policy has one or more constraints, each an aggregation or a path constraint. Path constraints have a test on a destination prefix and a list of regular expressions describing network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to that network. Values \In\ and \Out\ will match any internal and external location respectively. There are two types of predicates, as shown. A test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, $P$) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

@@

  \vspace*{1em}
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    % \inference{0}{L$(j,k)$}{$e_i > 0$}{S$(min(j*k,g),1)$}{I-out3}{$e_1$}{$e_2$}
    \inferencetwo{0}
    \node[anchor=west] at (1,2.2) {where $g \geq n-\dfrac{(m-j)*e_1}{e_2}$};
    \inferencelocal{6.5}{.4}
    \inferenceglobal{12}{.4}
  \end{tikzpicture}
  \end{minipage}

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we cannot exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. 
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top. 
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

Consider rule \infrule{I-out2}, and consider any node in the role $n$. There are $e_2$ incoming edges to that node. Due to the fact that $A_m(j,k)$, we know that at least j of those $e_2$ edges are connected to nodes with disjoint paths from the origin. Hence we infer $A_n(1,min(j,e_2))$.

%The rule \infrule{I-out2} follows a similar line of reasoning. It says that if any group of size $j$ with $k$ disjoint paths to each is reachable in the bottom role, and there is at least one incoming edge for each concrete node in the top role, then there are $min(j,e_2)$ disjoint paths to any single concrete node in the top role. This is because we can choose the $j$ nodes on the bottom for which we want to find disjoint paths, and then use each of the $e_2$ disjoint incoming edges for the particular node at the top connecting it to the bottom.

The two rules \infrule{I-mesh1} and \infrule{I-mesh2} handle the case where there is a full mesh between the two roles. This happens when the number of outgoing edges ($e_1$) from nodes in role $m$ equals the number of nodes on top ($n$). \infrule{I-mesh1} says that we can find disjoint paths to each node in the top role restricted to the number of disjoint paths we started with. \infrule{I-mesh2} uses the fact that each node in the top role is connected to each node in the bottom role to infer that there can be $j$ disjoint paths to any single node in the top role.

The annotation $mincut(X)$ appearing on an edge is an assertion about the fault tolerance between nodes in two different roles. The rule \infrule{I-mincut} uses such assertions. To each node in role $n$, from a node in role $m$, we can construct at least the minimum of $X$ and $k$ disjoint paths.

%The rule \infrule{I-mincut} can be thought of as an assertion about the fault tolerance between nodes in two roles. The rule applies to an edge that has been annotated with a $mincut(X)$ statement, which means that there are at least $X$ disjoint paths between any node in role $m$ and any node in role $n$.

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant $L_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least ``fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$), which each have $e_2$ incoming edges. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for datacenter topologies where routers in one tier of the datacenter often have a particular striping pattern with another tier.

Finally, rules \infrule{I-local} and \infrule{I-global} allow us to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from $L_m(j,k)$ to $L_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

@@

%The algorithm associates with each inference, the set of abstract locations used to derive the inference, \IE, an overapproximation of set of concrete locations used for the edge-disjoint paths. When applying an inference rule between an invariant at role $m$, with a set of abstract locations $C$, to learn an invariant at role $n$, we associate the new invariant with the set $C \cup \{ n \}$. 
%
%When, at a single PG node there is an inference $L_{X_1} \dots L_{X_{n-1}}$ A$_m(j,k)$ with location set $C$ and inference $L_{X_1} \dots L_{X_{n-1}}$ L$_m(j',k')$ with location set $D$, where $C \cap D = \emptyset$, we add a new inference with location set $C \cup D$: 
%
%$$L_{X_1} \dots L_{X_{n-1}} L_m(\min(j, j'), k + k')$$ 
%
%\noindent
%The idea is that, because non-overlapping abstract paths can never contain overlapping concrete paths, due to the graph homomorphism property, we can merge inferences about these disjoint paths in the abstraction.

@@

Note that, because each inference rule only applies to directed edges in the PG, the algorithm cannot make any inferences about connectivity from the AL role to the S role since there is no directed edge from AL to S. This restriction ensures that the analysis is policy-sensitive.

@@

In this case, the algorithm was able to infer that any node in the spine role has at least 1 disjoint path and any node in the TL node has at least 2 disjoint paths to a router in role TG. Note that, in this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint. In this case in particular, the bound on the fault tolerance between a ToR and spine router can be used to determine that an aggregation-induced black hole may arise at the border of the datacenter after a single failure in some topologies.

@@

To evaluate \sysname, we apply it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstracttions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

\subsection{Analysis precision}

\begin{figure}[t!]
  \begin{center}
      \begin{tabular}{| l | c | c | c | c | c | }
      \hline
      \textbf{Topology} & \textbf{Fixed} & \multicolumn{2}{|c|}{\textbf{Reachability}} & \multicolumn{2}{|c|}{\textbf{K-paths}} \\ \hline
      & & Some & All & Some & All \\
      & & Pairs & Pairs & Pairs & Pairs \\ \hline

      Fattree~\cite{fattree} & -- & \cmark & \cmark & \cmark & \xmark  \\ \hline
      Facebook~\cite{facebook-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10~\cite{f10-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      VL2~\cite{vl2-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      BCube~\cite{bcube} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
      DCell~\cite{dcell} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly~\cite{butterfly} & $n$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      HyperX~\cite{hyperx} & $L$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \vspace{-.6em}
  \caption{Fault-tolerance analysis precision for all topologies with shortest path routing.}
  \label{fig:analysis-precision}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \begin{center}
      \begin{tabular}{| l | c | c | c | c | }
      \hline
      \textbf{Clos Variant} & \multicolumn{2}{|c|}{\textbf{Reachability}} & \multicolumn{2}{|c|}{\textbf{K-paths}} \\ \hline
      & Some & All & Some & All \\
      & Pairs & Pairs & Pairs & Pairs \\ \hline

      Fattree & \cmark & \cmark & \cmark & \cmark  \\ \hline
      Facebook & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10 & \cmark & \cmark & \cmark & \cmark \\ \hline
      VL2 & \cmark & \cmark & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \vspace{-.6em}
  \caption{Fault-tolerance analysis precision for fat tree topologies with valley-free routing.}
  \label{fig:analysis-precision2}
\end{figure}

We evaluate the expressiveness of \sysname's abstractions and the precision of the fault-tolerance analysis on a range of different network topologies found both in the network literature and in production networks. The analysis is precise if it finds a tight lower-bound on the fault-tolerance (\IE, there is actually a concrete network with that fault tolerance).
For each topology in Figure~\ref{fig:analysis-precision} we define a suitable abstraction. Since many topologies have tunable parameters (\EG, the recursion depth $k$ for the DCell topology), we list all such parameters that are fixed for the abstraction. For all topologies (Figure~\ref{fig:analysis-precision}), we implement a shortest path routing policy. Additionally, for clos-style topologies, we implement a valley-free routing policy (Figure~\ref{fig:analysis-precision2}) for comparison. In each case, we record the precision of the analysis in terms of whether or not it is able to accurately prove reachability and/or the correct number of disjoint paths between some or all pairs of nodes from a fixed source.

\para{Fat Tree Topologies}

The fat tree topology is a widely used datacenter topology. We look at the precision of our analysis on four fat tree variants used in practice: a standard fat tree~\cite{fattree}, the Facebook fat tree~\cite{facebook-fattree}, the F10 fault tolerant fat tree~\cite{f10-fattree}, and the VL2 topology~\cite{vl2-fattree}. For each fat tree variant, we use a tiered abstraction similar to that in our previous example. Each fat tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion. For all fat tree variants, the analysis correctly computes the number of ToR to ToR disjoint paths, but can lose precision between ToR to spine routers due to valley paths. For a valley-free routing policy, the analysis is precise for ToR to spine connectivity.

\para{Recursive Topologies}

We also model several recursive topologies. These include the BCube~\cite{bcube} and DCell~\cite{dcell} topologies. Each topology includes a recursion depth parameter ($k$), which we fix. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but not from increasing the recursion depth.
For BCube, we model each tier of the datacenter as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.
The analysis is able to accurately determine reachability (i.e., at least 1 disjoint path) for each recursive topology, but is overly conservative in determining k-disjoint paths. For the BCube topology, it is able to to determine the correct number of disjoint paths between some pairs of nodes, but not all.

\para{Hypercube Topologies}

Hypercube variants can be used as an alternative to Clos-style topologies for networks with port density routers. The HyperX~\cite{hyperx} topology generalizes the hypercube and butterfly topologies and includes parameters $L$ for the lattice dimension of the network, and $S_i$ for the node multiplicity of each dimension $i$.
%$K_i$ for the bandwidth of links for each dimension,
%and $T$ for the number of terminal nodes (servers) connected to each router.
For a fixed number of dimensions $L$, we abstract each full mesh of $S_L$ nodes into its own abstract node. Nodes in dimension $S_{x-1}$ are abstracted using pods of abstract nodes from dimension $S_x$. Figure~\ref{fig:example-abstractions} shows an example for $L=2$.
For all hypercube variants, the fault-tolerance analysis is precise.

>>>>

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Example of a small configuration component of a router template for the datacenter spine role.}
  \label{fig:templates}
  \vspace{-1em}
\end{figure}

@@

%While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they provide no guarantees
As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. At the same time, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

\sysname takes as input the network's abstract topology and its policy. The policy has two components: the routing policy describes how traffic should flow, and the fault-tolerance policy describes how many simultaneous failures the network can withstand without loosing connectivity.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. The second line does the same for all local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the datacenter:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the variable $\CD{\$LP}$ and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the datacenter) at each hop of the path.
%
The constraint to prevent "valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each datacenter level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the datacenter through a peer.

@@

The first step for the \sysname compiler is to build the product graph (PG), a data structure that is amenable to joint analysis of the topology and routing policy. As a precursor to this step, we convert the routing policy syntax in \S\ref{sec:overview} to a core language based on regular expressions, shown in Figure~\ref{fig:syntax}.
%
A policy has one or more constraints, each an aggregation or a path constraint. Path constraints have a test on a destination prefix and a list of regular expressions describing network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to that network. Values \In~ and \Out, will match any internal and external location respectively. There are two types of predicates, as shown. A test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, $P$) where $G' = (V',E')$ is has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

@@

  \vspace*{1em}
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    % \inference{0}{L$(j,k)$}{$e_i > 0$}{S$(min(j*k,g),1)$}{I-out3}{$e_1$}{$e_2$}
    \inferencetwo{0}
    \node[anchor=west] at (1,2.2) {where $g=n-\dfrac{(m-j)*e_1}{e_2}$};
    \inferencelocal{6.5}{.4}
    \inferenceglobal{12}{.4}
  \end{tikzpicture}
  \end{minipage}

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since adding disjoint edges between the nodes keeps the existing paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.

The rule \infrule{I-out2} follows a similar line of reasoning. It says that if any group of size $j$ with $k$ disjoint paths to each is reachable in the bottom role, and there is at least one incoming edge for each concrete node in the top role, then there are $min(j,e_2)$ disjoint paths to any single concrete node in the top role. This is because we can choose the $j$ nodes on the bottom for which we want to find disjoint paths, and then use each of the $e_2$ disjoint incoming edges for the particular node at the top connecting it to the bottom.

The two rules \infrule{I-mesh1} and \infrule{I-mesh2} handle the case where there is a full mesh between the two roles. This happens when the number of outgoing edges from the bottom $e_1$ is equal to the number of nodes on the top $n$. The first case for \infrule{I-mesh1} says that we can find disjoint paths to each node in the top role restricted to the number of disjoint paths we started with. The second case for \infrule{I-mesh2} uses the fact that each node in the top role is connected to each node in the bottom role to infer that there can be $j$ disjoint paths to any single node in the top role.

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant L$_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least "fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

The second part of the rule allows uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for datacenter topologies where routers in one tier of the datacenter often have a particular striping pattern with another tier.

The rule \infrule{I-mincut} can be thought of as an assertion about the fault tolerance between nodes in two roles. The rule applies to an edge that has been annotated with a $mincut(X)$ statement, which means that there are at least $X$ disjoint paths between any node in role $m$ and any node in role $n$.

Finally, the \infrule{I-local} and \infrule{I-global} rules allow us to lift each of the other rules to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from L$_m(j,k)$ to L$_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

@@

The algorithm associates with each inference, the set of abstract locations used to derive the inference, \IE, an overapproximation of set of concrete locations used for the edge-disjoint paths. When applying an inference rule between an invariant at role $m$, with a set of abstract locations $C$, to learn an invariant at role $n$, we associate the new invariant with the set $C \cup \{ n \}$.
%
When, at a single PG node there is an inference $L_{X_1} \dots L_{X_{n-1}}$ A$_m(j,k)$ with location set $C$ and inference $L_{X_1} \dots L_{X_{n-1}}$ L$_m(j',k')$ with location set $D$, where $C \cap D = \emptyset$, we add a new inference with location set $C \cup D$:
%
$$L_{X_1} \dots L_{X_{n-1}} L_m(\min(j, j'), k + k')$$
%
\noindent
The idea is that, because non-overlapping abstract paths can never contain overlapping concrete paths, due to the graph homomorphism property, we can merge inferences about these disjoint paths in the abstraction.

@@

Note that, because each inference rule only applies to directed edges in the PG, the algorithm can not make any inferences about connectivity from the AL role to the S role since there is no directed edge between them. This restriction ensures that the analysis is policy-sensitive.

@@

In this case, the algorithm was able to infer that any node in the spine role has at least 1 disjoint path and any node in the TL node has at least 2 disjoint paths to a router in role TG. Note that, in this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global tor from a spine router due to the valleyfree constraint. In this case in particular, the bound on the fault tolerance between a tor and spine router can be used to determine that an aggregation-induced black hole may arise at the border of the datacenter after a single failures in some topologies.

@@

We evaluate \sysname by applying it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstractions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

\subsection{Expressiveness and precision}

\begin{figure}[t!]
  \begin{center}
      \begin{tabular}{| l | c| c | c | c | c | }
      \hline
       & \textbf{Fixed} & \multicolumn{2}{|c|}{\textbf{Reachability}} & \multicolumn{2}{|c|}{\textbf{K-paths}} \\ \hline
      & & Some & All & Some & All \\
      & & Pairs & Pairs & Pairs & Pairs \\ \hline

      \multicolumn{6}{l}{\textbf{Tree-based topologies, valley-free routing}} \\ \hline

      Fattree~\cite{fattree} & -- & \cmark & \cmark & \cmark & \cmark  \\ \hline
      Facebook~\cite{facebook-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10~\cite{f10-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      VL2~\cite{vl2-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
%      \end{tabular}
%  \end{center}
%  \vspace{-.6em}
%  \caption{Expressiveness and analysis precision for fat tree topologies with valley-free routing policy.}
%  \label{fig:analysis-precision2}
%\end{figure}
%
%\begin{figure}[t!]
%  \begin{center}
%      \begin{tabular}{| l | c | c | c | c | c | }
%      \hline
%      \textbf{Topology} & \textbf{Fixed} & \multicolumn{2}{|c|}{\textbf{Reachability}} & \multicolumn{2}{|c|}{\textbf{K-paths}} \\ \hline
%      & & Some & All & Some & All \\
%      & & Pairs & Pairs & Pairs & Pairs \\ \hline

      \multicolumn{6}{l}{\textbf{All topologies, shortest-path routing}} \\ \hline

      Fattree~\cite{fattree} & -- & \cmark & \cmark & \cmark & \xmark  \\ \hline
      Facebook~\cite{facebook-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10~\cite{f10-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      VL2~\cite{vl2-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      BCube~\cite{bcube} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
      DCell~\cite{dcell} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly~\cite{butterfly} & $n$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      HyperX~\cite{hyperx} & $L$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \vspace{-.6em}
  \caption{Expressiveness and precision of \sysname.}
  \label{fig:analysis-precision}
  \vspace{-1em}
\end{figure}

We evaluate the expressiveness of \sysname's topology abstractions and the precision of its fault-tolerance analysis on a range of network topologies found in production networks and in the networking literature. We characterize expressiveness by checking if our abstractions allow the topologies to evolve arbitrarily or certain aspects must be fixed (i.e., cannot be symbolic). We measure precision by checking if we find a tight lower-bound on fault-tolerance (\IE, there exists a concrete network with that fault-tolerance level).

The top part of Figure~\ref{fig:analysis-precision} shows the results for common real-world settings: tree-based topologies coupled with valley-free routing.
We consider four variants of tree topologies: a standard fat tree~\cite{fattree}, the Facebook fat tree~\cite{facebook-fattree}, the F10 fault-tolerant fat tree~\cite{f10-fattree}, and VL2~\cite{vl2-fattree}. These variants differ in the number of tiers and the connectivity pattern between roles. For each, we use a tiered abstraction similar to that in our example (\S\ref{sec:overview}) and parameterize over the number of pods, which can be scaled to allow expansion. We report precision of both analyzing reachability and computing the number of disjoint paths, and we report if \sysname is precise for all pairs of abstract nodes or only some of them. %"All pairs" precision implies "some pairs" precision, but not the other way around.

Our results are encouraging for these realistic settings. Our abstractions are perfectly expressive for all tree-based topologies---we did not have to fix any aspect of their structure---and the analysis is precise in all cases.

%For each topology in Figure~\ref{fig:analysis-precision} we define a suitable abstraction. Since many topologies have tunable parameters (\EG, the recursion depth $k$ for the DCell topology), we list all such parameters that are fixed for the abstraction. For all topologies (Figure~\ref{fig:analysis-precision}), we implement a shortest path routing policy. Additionally, for clos-style topologies, we implement a valley-free routing policy (Figure~\ref{fig:analysis-precision2}) for comparison. In each case, we record the precision of the analysis in terms of whether or not it is able to accurately prove reachability and/or the correct number of disjoint paths between some or all pairs of nodes from a fixed source.

%\para{Fat Tree Topologies}
%
%The fat tree topology is a widely used datacenter topology. We look at the precision of our analysis on four fat tree variants used in practice: a standard fat tree~\cite{fattree}, the Facebook fat tree~\cite{facebook-fattree}, the F10 fault tolerant fat tree~\cite{f10-fattree}, and the VL2 topology~\cite{vl2-fattree}. For each fat tree variant, we use a tiered abstraction similar to that in our previous example. Each fat tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion.

To stress our abstractions and analysis, we consider additional topologies from the literature.

\para{Recursive Topologies}
These topologies include BCube~\cite{bcube} and DCell~\cite{dcell}. Each topology includes a recursion depth parameter ($k$), which we fixed while abstracting them. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but does not allow changing the recursion depth dynamically. 
For BCube, we model each tier of the datacenter as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.

\para{Hypercube Topologies}
Hypercube variants can be used as an alternative to Clos-style topologies for networks with port density routers. The HyperX~\cite{hyperx} topology generalizes the hypercube and butterfly topologies and includes parameters $L$ for the lattice dimension of the network, and $S_i$ for the node multiplicity of each dimension $i$.
%$K_i$ for the bandwidth of links for each dimension,
%and $T$ for the number of terminal nodes (servers) connected to each router.
For a fixed number of dimensions $L$, we abstract each full mesh of $S_L$ nodes into its own abstract node. Nodes in dimension $S_{x-1}$ are abstracted using pods of abstract nodes from dimension $S_x$. Figure~\ref{fig:example-abstractions} shows an example for $L=2$.
%For all hypercube variants, the fault-tolerance analysis is precise.
